---
title: "Team 5 Notebook: King County Housing"
author: Fucheng Yao, Limei Huang, Eman Nagib, Kwangwoo Kim, Huaiping Wang
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggthemes)
library(glmnet)
```

## Problem: 
> What is the housing price given a certain housing criteria?

### Project impact:
>For owners - understand the value of your property relative to other properties that offer the same features
>
>For potential investors - assert if the house’s value is reasonable etc.
>
>For potential buyers/brokers - comparing properties and finding the best value proposition for a certain criteria


## Data overview:
>Data source: Kaggle - https://www.kaggle.com/harlfoxem/housesalesprediction
>
>house_id - Unique ID for each home sold
>
>Year_sold, month_sold, day_sold - Date of the home sale
>
>price - Price of each home sold
>
>bedrooms - Number of bedrooms
>
>bathrooms - Number of bathrooms, where .5 accounts for a room with a toilet but no shower
>
>sqft_living - Square footage of the apartments interior living space
>
>sqft_lot - Square footage of the land space
>
>floors - Number of floors
>
>waterfront - A dummy variable for whether the apartment was overlooking the waterfront or not
>
>view - An index from 0 to 4 of how good the view of the property was
>
>condition - An index from 1 to 5 on the condition of the apartment
>
>grade - An index from 1 to 13, where 1-3 falls short of building construction and design, 7 has an average level of construction and design, and 11-13 have a high quality level of construction and design.
>
>sqft_above - The square footage of the interior housing space that is above ground level
>
>sqft_basement - The square footage of the interior housing space that is below ground level
>
>yr_built - The year the house was initially built
>
>yr_renovated - The year of the house’s last renovation
>
>zipcode - What zip code area the house is in
>
>lat - Latitude
>
>long - Longitude
>
>sqft_living15 - The square footage of interior housing living space for the nearest 15 neighbors
>
>sqft_lot15 - The square footage of the land lots of the nearest 15 neighbors

```{r}

```

